Distributed Multi-Heuristic A* (HAMSTAR)

### 2014-12-11 paper.tex
Updated paper stuff

We investigated other frameworks designed for iterative algorithms on a stable dataset such as Spark and GraphLab. However, most of these frameworks were designed for chunking up a large dataset upon which a single operation is performed. Our application's needs differed from this model in that the dataset being operated on (the graph being explored) was shared by all workers, and the workers themselves were performing differing operations on subsets of the same data.

We chose MPI, despite its complexity and low level of abstraction, because our algorithm requires extremely fast and efficient communication between the anchor and the child machines. Moreover, since the entire state must be replicated across all machines, the necessary communication was relatively straightforward. Finally, we found detailed documentation on MPI, explaining the low-level workings of each command, much more readily accessible than comparable documentation on Spark or GraphLab, which seemed more focused on abstracting the details of the system away from the programmer. Given our emphasis on speed, MPI seemed like an appropriate choice. We ultimately used the OpenMPI implementation of the MPI standard.

@@

We considered two communication models for our algorithm. The first one has the anchor periodically send each child a table of states. The table contains information for new states that were either received by the anchor, or calculated by the anchor, since the last time the anchor communicated with this child. Each child frequently checks for incoming messages using the \emph{MPI\_Iprobe} command; if a message is detected, the child receives the incoming data using \emph{MPI\_Recv} and responds with its own table of states. The anchor sends messages to each child one at a time, moving on to the next child when a response is received.

The second communication model involves a period broadcast using the \emph{MPI\_Bcast} command. After a pre-determined number of iterations, all child machines send their new states to the anchor using \emph{MPI\_Gather}. The anchor sifts through these responses, and keeps only the best $g$-value for each state. Every state that has had a new best $g$-value found is then broadcast to all the child machines.

The one-on-one communication model has the benefit that it does not require all machines to cease work while communication happens; only the anchor and the single child machine need to participate in the communication. The downside is that the anchor must then do $O(N)$ communications, where $N$ is the number of children, and the time it takes to propogate the new state tables across the entire system is also $O(N)$. Broadcasts, on the other hand, require a simultaneous call from all machines. This generally requires more overhead, because it is unlikely that all machines will call the command at precisely the same time. However, once called, the Broadcast command allows the anchor to only send the data to a few machines in the network. These machines, in turn, forward the data to other machines, and so on. Thus, the anchor does not spend any more time communicating than any other machine in the system, and the new state table is propagated in $O(\log(N))$ time.

>>>>

We investigated other frameworks designed for iterative algorithms on a stable dataset such as Spark and GraphLab. However, most of these frameworks were designed for chunking up a large dataset and assigning pieces to workers. Our application's needs differed from this model in that the dataset being operated on (the graph being explored) was shared by all workers, and the workers themselves were performing differing operations on subsets of the same data.

We chose MPI, despite its complexity and low level of abstraction, because our algorithm requires extremely fast and efficient communication between the anchor and the child machines. Moreover, since the entire state must be replicated across all machines, the necessary communication was relatively straightforward. Finally, we found that detailed documentation on MPI, explaining the low-level workings of each command, was much more readily accessible than comparable documentation on Spark or GraphLab, which seemed more focused on abstracting the details of the system away from the programmer. Given our emphasis on speed, MPI seemed like an appropriate choice. We ultimately used the OpenMPI implementation of the MPI standard.

@@

We considered two communication models for our algorithm. The first one, which we call the one-on-one model, has the anchor periodically send all children a table of states. The table contains information for new states that were either received by the anchor, or calculated by the anchor, since the last time the anchor communicated with this child. Each child frequently checks for incoming messages using the \emph{MPI\_Iprobe} command; if a message is detected, the child receives the incoming data using \emph{MPI\_Recv} and responds with its own table of states. The anchor sends messages to each child one at a time, moving on to the next child when a response is received. Once the data has been sent to all children, the anchor will run its own iterations until the next cycle should begin.

The second communication model involves a periodic broadcast using the \emph{MPI\_Bcast} command. After a pre-determined number of iterations, all child machines send their new states to the anchor using \emph{MPI\_Gather}. The anchor sifts through these responses, and keeps only the best $g$-value for each state. Every state that has had a new best $g$-value found is then broadcast to all the child machines.

The one-on-one communication model has the benefit that it does not require all machines to cease work while communication happens; only the anchor and a single child machine need to participate in the communication. The downside is that the anchor must then do $O(N)$ communications, where $N$ is the number of children, and the time it takes to propogate the new state tables across the entire system is also $O(N)$. Broadcasts, on the other hand, require a simultaneous call from all machines. This generally requires more overhead, because it is unlikely that all machines will call the command at precisely the same time. However, once called, the Broadcast command allows the anchor to only send the data to a few machines in the network. These machines, in turn, forward the data to other machines, and so on. Thus, the anchor does not spend any more time communicating than any other machine in the system, and the new state table is propagated in $O(\log(N))$ time.
